Clustering Constraints on the Relative Sizes of Central and Satellite Galaxies

### 2017-11-09 paper/galsize_paper.tex
edited S 5, including 5.2

Recall that in the $\rvir$-based size model, galaxy size is proportional to the virial radius of host halo at the redshift at the time the halo attained its maximum mass. Thus, we can expect that in this model satellite galaxies will have sizes smaller than sizes of central galaxies hosted by halos of similar stellar mass.  Indeed, Figure \ref{fig:censatsizehist} shows the $\rhalf$ distributions of central, satellite, and ``splashback'' central galaxies hosted by halos of the same mass $\mhalo\equiv\mpeak\approx10^{12}\msun.$ A ``splashback central"  is defined as a present-day central galaxy that used to be a satellite, i.e., its main progenitor halo passed inside the virial radius of a larger halo at some point in its past history. On the other hand, we define a ``true central" as a galaxy that has never been a satellite. The figure shows that there is a clear trend of sizes decreasing from the central to splashback and to satellite galaxies.

@@

While we consider such an explanation plausible, and even likely, a conclusive confirmation requires proper forward modeling of size and color jointly. Such an effort is well-motivated, but beyond the scope of the present work.

\subsection{Broader Implications}
\label{subsec:broader_implications}

Physical explanations for differences between the sizes of centrals and satellites fall into two categories, depending on whether or not the processes involved are post-infall and satellite-specific. In the satellite mass loss model, satellites evolve exactly as centrals prior to the time of infall; this model assumes that {\em all} satellite size truncation occurs after the time of infall. The $\rvir-$based model assumes the opposite: namely that the size of {\em all} galaxies is statistically set by the $\rmpeak,$ the physical size of the virial radius at the time at which $\mhalo$ stops growing. This assumption is the natural extension of the abundance matching approximation that stellar mass largely stops growing at this time (including in-situ and ex-situ growth). The halos of typical satellites reach $\mpeak$ far outside the virial radius of their ultimate host halo \citep{behroozi_etal14}, and so in the $\rvir-$based model the size differences between satellites and centrals is largely in place prior to the time of satellite infall.

The clustering ratios in Figure \ref{fig:mstarmodelclustering} show that the impact of physically motivated levels of satellite mass loss likely cannot explain the $\rhalf-$dependent clustering seen in SDSS. The success of the $\rvir-$based model, together with the failure of the satellite mass loss model, suggests that the relative size of central and satellite galaxies is largely driven by the {\rm pre-infall} evolution of the satellite.

@@

Numerous previous analyses have employed the \citet{yang_etal05b} group catalog to ``directly" study the relationship between galaxy mass, size, and environment. For example, by treating observed groups as genuine dark matter halos, several analyses have found that the $\mstar-\rhalf$ relation of early-type galaxies exhibits weak, if any, environmental dependence \citep{weinmann_etal08,huertas_company_etal13b,shankar_etal14}. Meanwhile, \citet{weinmann_etal08} additionally find that for late-type galaxies, there is a strong correlation between disk size and central/satellite designation. Qualitatively, our clustering measurements shown in Figure \ref{fig:colorclustering} are in good agreement with these findings: once galaxies have been morphology-selected, only the clustering of disk shows significant dependence upon $\rhalf.$ Our clustering results are also consistent with the disk model presented in \citet{dutton_etal08,dutton_etal10}, which predicts that at fixed stellar mass, there is weak, positive correlation between disk size and halo mass.

>>>>

Recall that in the $\rvir$-based size model, galaxy size is proportional to the virial radius of host halo at the redshift  when the halo halo attained its maximum mass. Thus, we can expect that in this model satellite galaxies will have sizes smaller than sizes of central galaxies hosted by halos of similar stellar mass.  Indeed, Figure \ref{fig:censatsizehist} shows the $\rhalf$ distributions of central, satellite, and ``splashback'' central galaxies hosted by halos of the same mass $\mhalo\equiv\mpeak\approx10^{12}\msun.$ A ``splashback central"  is defined as a present-day central galaxy that used to be a satellite, i.e., its main progenitor halo passed inside the virial radius of a larger halo at some point in its past history. On the other hand, we define a ``true central" as a galaxy that has never been a satellite. The figure shows that there is a clear trend of sizes decreasing from the central to splashback and to satellite galaxies.

@@

While we consider such an explanation plausible, and even likely, a conclusive confirmation requires proper forward modeling of size and color jointly. Such an effort is strongly motivated by our results, but is beyond the scope of the present work.

%-------------------------------------------------
\subsection{Broader Implications}
\label{subsec:broader_implications}
%-------------------------------------------------

The success of the $\rvir$-based model in explaining the size dependence of galaxy clustering implies that sizes of {\it all} galaxies,  red, blue, spheroid, disks, satellites and central, are, on average, proportional to the $\rmpeak,$ the physical size of the virial radius at the time at which $\mhalo$ stops growing. Although correlation is not necessarily causation, such universal and simple dependence poses a useful challenge to galaxy formation models, as it indicates tight connection betw
een growth of galaxies and their host halos.

The $\rvir$-based size model also assumes that after halo stops growing the size remains approximately constant and this assumption is consistent with observed size dependence of galaxy clustering in SDSS.  The halos of typical satellites reach $\mpeak$ far outside the virial radius of their ultimate host halo \citep{behroozi_etal14}, and so in the $\rvir-$based model the size difference between satellites and centrals is largely in place prior to the time of satellite infall. 
Thus, as noted above, halos of central and satellite galaxies of the same $\mstar$  differ in epochs when their halo stopped growing (i.e., $\zpeak$) and this difference explains their different median sizes.

The difference also explains the relative trends of $\rhalf-\mstar$ relations for red and blue, or disk and spheroid galaxies.
Both types of galaxies exhibit weak dependence of $\rhalf$ on $\mstar$ at $\mstar\lesssim 5\times 10^{10}\ M_\odot$ and 
at these masses $\rhalf$ of disks is a factor of two larger than those of spheroids. At larger masses, the sizes of galaxies of both types show stronger dependence on $\mstar$ and the difference in median size of disk and spheroid galaxies decreases \citep[e.g.,][]{bernardi_etal14}.

In the $\rvir$-based size model, the weak dependence of $\rhalf$ on $\mstar$ for low mass galaxies is due to steep $\mstar-\mvir$ relation. For example, if $\mstar\propto\mvir^\alpha\propto\rvir^{3\alpha}$, then $\rhalf\propto \rvir\propto \mstar^{1/(3\alpha)}$. Given that $\alpha\sim 1-3$ is inferred for the $\mstar-\mvir$ relation at low masses \citep{kravtsov10,behroozi13_smhm,kravtsov_etal14}, the $\rhalf-\mstar$ relation in this regime is expected to be shallow. 
For galaxies of $\mstar\gtrsim 5\times 10^{10}\ M_\odot$, $\alpha\approx 0.3$ \citep[e.g.,][]{kravtsov_etal14} and thus 
the slope of the $\rhalf-\mstar$ relation is approximately linear. These trends of $\rhalf$ with $\mstar$ have a direct relation to the trends of surface brightness $\mu_\star=\mstar/(\pi\rhalf^2)$ with stellar mass.

The difference in median sizes of disk and spheroidal galaxies at $\mstar\lesssim 10^{11}\ M_\odot$ is explained by the fact that sizes of satellite galaxies are systematically smaller than sizes of central galaxies of the same $\mstar$ and   
the satellite fraction is much higher among spheroidal galaxies compared to the disk galaxies. For example, analyses of \citet{rodriguez_puebla_etal15} show that in this regime satellite fraction of blue galaxies is only $\approx 10-20\%$, while
the satellite fraction of red galaxies is $\gtrsim 30-50\%$ increasing with decreasing stellar mass. 
As $\mstar$ increases, satellite fractions of both blue and red galaxies drop and at $\mstar\gtrsim 10^{11}\ M_\odot$ vast majority of blue and red galaxies are centrals. If these galaxies occupy similar halo masses at the same $\mstar$, $\rvir$-based size model predicts that  the difference in sizes of red and blue galaxies should disappear at these masses, as is indeed observed.

A testable prediction of this picture is that when 3D (de-projected) sizes of only central red and blue galaxies of the same $\mstar$ are compared they should be much closer, than the sizes of the overall red and blue populations.

@@

Numerous previous analyses have employed the \citet{yang_etal05b} group catalog to ``directly" study the relationship between galaxy mass, size, and environment. For example, by treating observed groups as genuine dark matter halos, several analyses have found that the $\mstar-\rhalf$ relation of early-type galaxies exhibits weak, if any, environmental dependence \citep{weinmann_etal08,huertas_company_etal13b,shankar_etal14}. Meanwhile, \citet{weinmann_etal08} additionally find that for late-type galaxies, there is a strong correlation between disk size and central/satellite designation. Qualitatively, our clustering measurements shown in Figure \ref{fig:colorclustering} are in good agreement with these findings: once galaxies have been color-selected, only the clustering of blue galaxies shows significant dependence upon $\rhalf.$ Our clustering results are also consistent with the disk model presented in \citet{dutton_etal08,dutton_etal10}, which predicts that at fixed stellar mass, there is weak, positive correlation between disk size and halo mass.
